---
title: "p8105_hw2_dc3964"
author: "Rene Chi"  
date: "2025-09-24"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1

```{r, message = FALSE}
## Problem 1
### loading packages
library(tidyverse)
library(lubridate)

### tidying pol-month
pols_month =
  read_csv("./data/pols-month.csv") |> 
  separate(mon, c("year", "month", "day")) |> 
  mutate(month = month.abb[as.integer(month)]) |> 
  mutate(president = case_when(
    prez_gop == 1 ~ "gop",
    prez_dem == 1 ~ "dem",
    TRUE          ~ NA_character_
  )) |> 
  select(-day, -prez_gop, -prez_dem)

### tidying snp
snp = 
  read_csv("./data/snp.csv") |> 
  separate(date, c("month", "day", "year")) |> 
  select(year, month, everything()) |> 
  select(-day) |> 
  mutate(month = month.abb[as.integer(month)]) |> 
  ## changed format from 2 digit to 4 digit 
  mutate(year = if_else(as.integer(year) <= 15,
                          2000 + as.integer(year),
                          1900 + as.integer(year))) |> 
  ## changed format from integer to character
  mutate(year = as.character(year))

### tidying unemployment
unemployment =
  read_csv("./data/unemployment.csv") |> 
  rename(year = Year) |> 
  mutate(year = as.character(year)) |> 
  pivot_longer(Jan:Dec, names_to = "month", values_to = "unemployment_rate")

### Merging all datasets
partial_complete = left_join(pols_month, snp) 
complete_data_p1 = inner_join(partial_complete, unemployment)
```

Pols-month has 822 rows and 9 variables, with the years ranged from 1947 to 2015. Key variables being year and month for future data merge.

Snp has 787 rows and 3 variables, with the years ranged from 1950 to 2015. Key variables being year and month for future data merge.

Unemployment has 816 rows and 3 variables, with the years ranged from 1948 to 2015. Key variables being year and month for future data merge.

The merged data set has 822 rows and 11 columns, with the years ranged from 1947 to 2015. Key variables being year and month as I used them to merge all of the data sets.

## Problem 2

TBD

```{r}
#TBD
```

## Problem 3

TBD

```{r}
#TBD
```
